---
title: "p8105_hw1_zl3119"
author: "Zheyan"
date: "2021/9/21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# English environment
Sys.setenv(LANGUAGE = "en")
# import packages
library(tidyverse)
```

# Problem 1

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

## Create a DataFrame

Create a DataFrame comprised of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r Create DataFrame}
# Generate a DataFrame with size 10 from a standard Normal distribution
rnorm_df = tibble(
  rnum = rnorm(10, sd = 1)
) 

final_df = rnorm_df %>% 
  # a logical vector indicating whether elements of the sample are greater than 0
  mutate(if_positive = ifelse(rnum > 0, TRUE, FALSE)) %>%
  # a character vector of length 10
  mutate(animals = sample(c('cat', 'dog', 'lion', 'panda', 'elephant'),size = 10, replace = TRUE)) %>%
  # factor vector of length 10, with 3 different factor “levels”
  mutate(value_size = factor(ifelse(rnum > - 0.5, ifelse(rnum < 0.5, 'medium', 'big'), 'small'))) 
  
```

## Take the mean of each variable 

```{r Take mean}
mean(final_df$rnum)
mean(final_df$if_positive)
mean(final_df$animals)
mean(final_df$value_size)
```
It only works for numerical and bool variables.

## Convert variables to number
Use "as.numeric" function to the logical, character, and factor variables

```{r Convert to number by using as.numeric, echo = FALSE}
#as.numeric(final_df$rnum)
print('logical variable')
as.numeric(final_df$if_positive)
print('character variable')
as.numeric(final_df$animals)
print('factor variables')
as.numeric(final_df$value_size)
```
logical variable will be transformed into 0 and 1. Character variable will be forced to become NA. Also, factors will be transformed into levels. And we can take the mean of levels as below.
```{r}
mean(as.numeric(final_df$value_size))
```

























