---
title: "p8105_hw1_zl3119"
author: "Zheyan"
date: "2021/9/21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# English environment
Sys.setenv(LANGUAGE = "en")
# import packages
library(tidyverse)
```

# Problem 1

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

## Create a DataFrame

Create a DataFrame comprised of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r Create DataFrame}
# Generate a DataFrame with size 10 from a standard Normal distribution
rnorm_df = tibble(
  rnum = rnorm(10, sd = 1)
) 

final_df = rnorm_df %>% 
  # a logical vector indicating whether elements of the sample are greater than 0
  mutate(if_positive = ifelse(rnum > 0, TRUE, FALSE)) %>%
  # a character vector of length 10
  mutate(animals = sample(c('cat', 'dog', 'lion', 'panda', 'elephant'),size = 10, replace = TRUE)) %>%
  # factor vector of length 10, with 3 different factor “levels”
  mutate(value_size = factor(ifelse(rnum > - 0.5, ifelse(rnum < 0.5, 'medium', 'big'), 'small'))) 
  
```

## Take the mean of each variable 

```{r Take mean}
mean(final_df$rnum)
mean(final_df$if_positive)
mean(final_df$animals)
mean(final_df$value_size)
```
It only works for numerical and bool variables.

## Convert variables to number
Use "as.numeric" function to the logical, character, and factor variables

```{r Convert to number by using as.numeric, echo = FALSE}
#as.numeric(final_df$rnum)
print('logical variable')
as.numeric(final_df$if_positive)
print('character variable')
as.numeric(final_df$animals)
print('factor variables')
as.numeric(final_df$value_size)
```
logical variable will be transformed into 0 and 1. Character variable will be forced to become NA. Also, factors will be transformed into levels. And we can take the mean of levels as below.
```{r}
mean(as.numeric(final_df$value_size))
```

# Problem 2

```{r, include=FALSE}
# install and import packages
# install.packages("palmerpenguins")
library(palmerpenguins)
library(gridExtra)
```

## Descibe the penguins dataset
Write a short description of the penguins dataset (not the penguins_raw dataset) using inline R code, including:

* the data in this dataset, including names / values of important variables
* the size of the dataset (using nrow and ncol)
* the mean flipper length
```{r}
# read data
data("penguins", package = "palmerpenguins")
penguins
```

Brief introduction of the data.

### Size of the DataFrame

The penguins DataFrame contains `r nrow(penguins)` observations and `r ncol(penguins)` features.

### Variables

Use the "str" function to have a brief understanding of the data.

```{r}
str(penguins)
```

There are 3 factor variables and 5 numeric variables. And some of them contains missing value, the count of each missing value is listed below.

```{r}
apply(is.na(penguins), 2, FUN = sum)
```

**Factor Variables**

```{r}
factor_features = c('island', 'species', 'sex')
apply(penguins[,factor_features], 2, unique)
```

For 3 factor variables, island takes on 3 levels (Biscoe, Dream and Torgersen), species takes on 3 levels (Adelie, Chinstrap, Gentoo) and sex takes on Female and Male. 

**Numeric Variables**

For 5 numeric variables, the quantiles of them are listed below (Missing value unincluded).

```{r}
numercal_features = c('bill_length_mm', 'bill_depth_mm', 'flipper_length_mm', 'body_mass_g', 'year')
apply(penguins[, numercal_features], 2, quantile, na.rm = TRUE)
```

### Mean flipper_length_mm

The mean flipper_length_mm of all three penguins is **`r round(mean(penguins$flipper_length_mm, na.rm = TRUE),2)`**

Flipper length varys between each species, therefore we need to calculate mean flipper length of each species.

```{r groupby and calculate mean}
# remove missing values
penguins[complete.cases(penguins[, c('species','flipper_length_mm')]), ] %>%
  group_by(species) %>%
  summarise_at(vars(flipper_length_mm), list(avg = mean))
```

## Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x)

```{r scatterplot of flipper_length_mm (y) vs bill_length_mm (x)}
penguins_complete = penguins[complete.cases(penguins[, c('species','bill_length_mm','flipper_length_mm')]), ]
ggplot(penguins_complete,aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point()
```
Adelie is the smallest of three (smallest bill and flipper); Chinstrap has the longest bill; Gentoo has the biggest flipper.

```{r}
# Save the plot as pdf (scatter_plot_bill_flipper.pdf)
ggsave('scatter_plot_bill_flipper.pdf', height=4, width=4)
```



