---
title: "Visualization1"
author: "Zheyan"
date: "2021/10/5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

```{r Import libraries}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```


# Scatterplot

## Basic Scatterplot
tmax vs tmin

```{r pressure, echo=FALSE}
weather_df %>% ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

you can save ggplot
```{r}
ggp_tmax_tmin = 
  weather_df %>% ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_tmax_tmin
```

## Advancecd scatterplot

Add color 

```{r}
weather_df %>% ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

We can also make the point more transparent and add a smooth line in it

```{r}
# add color in ggplot, the color is global/public (in point and smooth line)
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE)
```

```{r}
# use facet to draw mutiple plots
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(.~name)
```

Lets make one more scatterplot
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name,size = prcp)) +
  geom_point() +
  facet_grid(.~name) + 
  geom_smooth(se = FALSE)

```

## Use data manipulations as part of this
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_point()
```

# Stacking geoms

Which geoms do you want

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth()
```

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_hex()
```

## Univariate plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(.~name)
```

Try other plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.7)
```

Still with tmax and name
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Some people like violin plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin()
```

Some linke ridges

```{r}
library(ggridges)
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges()
```

## Embedding plots

```{r trying to show on github, fig.width=7, fig.height=5}
weather_df %>% ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```













