---
title: "Visualization2"
author: "Zheyan"
date: "2021/10/11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
```

# Weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

head(weather_df)
```
# Lets plot!

## Label and title

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = 'Temperature Plot',
    x = 'Min Temperature',
    y = 'Max Temperature',
    caption = 'Data from rnoaa package; temperature in 2017'
  )
```

## Scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = 'Temperature Plot',
    x = 'Min Temperature',
    y = 'Max Temperature',
    caption = 'Data from rnoaa package; temperature in 2017'
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c('-15 C', '0 C', '15 C')
  ) +
  scale_y_continuous(
    trans = 'sqrt',
    position = 'right'
  )
```

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = 'Temperature Plot',
    x = 'Min Temperature',
    y = 'Max Temperature',
    caption = 'Data from rnoaa package; temperature in 2017'
  ) +
  scale_color_hue(
    name = 'Location',
    h = c(100,300)
  )
```

Also, we can do this with viridis package

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = 'Temperature Plot',
    x = 'Min Temperature',
    y = 'Max Temperature',
    caption = 'Data from rnoaa package; temperature in 2017'
  ) +
  viridis::scale_color_viridis(
    name = 'Location',
    discrete = TRUE)

```

## Themes

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = 'Temperature Plot',
    x = 'Min Temperature',
    y = 'Max Temperature',
    caption = 'Data from rnoaa package; temperature in 2017'
  ) +
  viridis::scale_color_viridis(
    name = 'Location',
    discrete = TRUE) +
  theme(legend.position = 'bottom')
```

Change the overall theme

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = 'Temperature Plot',
    x = 'Min Temperature',
    y = 'Max Temperature',
    caption = 'Data from rnoaa package; temperature in 2017'
  ) +
  viridis::scale_color_viridis(
    name = 'Location',
    discrete = TRUE) +
  # theme_minimal()
  theme_bw()+
  theme(legend.position = 'bottom')
```

## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in 'geom'

```{r}
central_park = 
  weather_df %>% 
  filter(name == 'CentralPark_NY')

waikiki = 
  weather_df %>% 
  filter(name == 'Waikiki_HA')

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point()+
  geom_line(data = central_park)
```

## 'patchwork'

remember faceting?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name) 
```

What happends when you want multipanel plots but can't facet ...?

```{r}
tmax_tmin_p = 
  weather_df %>% 
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .5) +
    theme(legend.position = 'none')

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5)  

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = 'none')

# NB
tmax_tmin_p + prcp_dens_p + tmax_date_p
tmax_tmin_p / (prcp_dens_p + tmax_date_p)
```


## Data manipulation

Control your factors.

```{r}
# fix alphabetly order of name
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c('Waikiki_HA'))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5) 
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
  filter(name == 'CentralPark_NY') %>% 
  pivot_longer(
    tmax:tmin,
    names_to = 'observation',
    values_to = 'temperatures'
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

add some facets

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = 'observation',
    values_to = 'temperatures'
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```


