---
title: "StringAndFactor"
author: "Zheyan"
date: "10/21/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(p8105.datasets)

library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# String and regex

Detect and select

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

string_vec[str_detect(string_vec, "jeff")]
```
Replace

```{r}
str_replace(string_vec, "jeff", "Jeff")
```

You can designate a list of characters that will count as a match.

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush")
```

You don’t have to list these; instead, you can provide a range of letters or numbers that count as a match.

```{r}
#sth starts with number
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

The character . matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Some characters are “special”. These include [ and ], ( and ), and .. If you want to search for these, you have to indicate they’re special using \. Unfortunately, \ is also special, so things get weird.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

# Factors

```{r}
factor_vec = factor(c('male', 'female', 'male', 'female'))

as.numeric(factor_vec)

factor_vec = fct_relevel(factor_vec, 'male')
```

## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)

table_marj
```

Clean Data

```{r}
data_marj = 
  table_marj %>% 
    select(-contains('P Value')) %>% 
    pivot_longer(
      -State,
      names_to = 'age_year',
      values_to = 'percent') %>% 
    separate(age_year, into = c('age', 'year'), sep = '\\(') %>% 
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = as.numeric(str_replace(percent, '[a-c]$', ''))
    ) %>% 
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

```

Do DataFrame stuff

```{r}
  
data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## NYC restaurants

```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec = 
  rest_inspec %>% 
  filter(
    str_detect(grade, '[ABC]'),
    !(boro == 'Missing')
  )

rest_inspec %>% 
  janitor::tabyl(boro, grade)

```

Focus on Pizza rating
```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```


That doesn’t look right – for sure there are more pizza place ratings than that! The problem is that the match in str_detect is case-sensitive until we tell it not to be:

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)
```
Might help to have things in a different order – maybe number of pizza places? Fortunately this can be done using fct_infreq.

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro),
         # How to change Manhatten into cities
         boro = fct_recode(boro, 'The City' = 'MANHATTAN')) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```

















