---
title: "Linear_models"
author: "Zheyan"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

# Model fitting

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)

nyc_airbnb
```

An good place to start is to consider price as an outcome that may depend on rating and borough. We fit that initial model in the following code.

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
summary(fit)
```

The lm function begins with the formula specification – outcome on the left of the ~ and predictors separated by + on the right. As we’ll see shortly, interactions between variables can be specified using *. You can also specify an intercept-only model (outcome ~ 1), a model with no intercept (outcome ~ 0 + ...), and a model using all available predictors (outcome ~ .).

R will treat categorical (factor) covariates appropriately and predictably: indicator variables are created for each non-reference category and included in your model, and the factor level is treated as the reference. As with ggplot, being careful with factors is therefore critical!

```{r}

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))


fit = lm(price ~ stars + borough, data = nyc_airbnb)
summary(fit)
```
If you want to present sth....


```{r}
fit %>%  
  broom::tidy() %>%
  mutate(term = str_replace(term, 'borough', 'Borough:')) %>% 
  select(term, estimate, p.value)
  
```

It’s important to note that changing reference categories won’t change “fit” or statistical sigificance, but can affect ease of interpretation.


# Diagnosis

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = resid)) +
  geom_density() +
  xlim(-250, 250)
```

## Interactions? Nesting?

Let's try a different 

```{r}
nyc_airbnb %>% 
  lm(price ~ stars * borough + room_type * borough, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

Try nesting

```{r}
nest_lm_res =
  nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(data, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(-data, -models) %>% 
  unnest(results)


nest_lm_res
```


Show model result of nesting data

```{r}
nest_lm_res %>% 
  select(borough, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, values_from = estimate) %>% 
  knitr::kable(digits = 3)
```

```{r}
manhattan_airbnb =
  nyc_airbnb %>% 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb %>% 
  nest(data = -neighborhood) %>% 
  mutate(
    models = map(data, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(-data, -models) %>% 
  unnest(results)

manhattan_nest_lm_res
```

Look at neighborhoods in Manhattan ...

```{r}
manhattan_nest_lm_res %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

manhattan_nest_lm_res %>% 
  distinct(term)

```

## sth beyond course

With this many factor levels, it really isn’t a good idea to fit models with main effects or interactions for each. Instead, you’d be best-off using a mixed model, with random intercepts and slopes for each neighborhood. 

```{r}
install.packages('broom.mixed')
manhattan_airbnb %>% 
  lme4::lmer(price ~ stars + room_type + (1 + room_type | neighborhood), data = .) %>% 
  broom.mixed::tidy()
```


# binary outcomes

data

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") %>% 
  filter(city == "Baltimore") %>% 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>% 
  select(resolved, victim_age, victim_race, victim_sex)

baltimore_df
```


model  

```{r}
fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

fit_logistic
```

```{r}
fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```

Give fitted result

```{r}
baltimore_df %>% 
  modelr::add_predictions(fit_logistic) %>% 
  mutate(fitted_prob = boot::inv.logit(pred))
```








