---
title: "p8105_hw2_zl3119"
author: "Zheyan"
date: "2021/9/30"
output: github_document
---

```{r import packages, include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem1
Read and clean the Mr. Trash Wheel sheet:

* specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

## Read and clean Mr.Trash Wheel Sheet

```{r Read and clean Mr.Trash Wheel,warning=FALSE}
TW_df_raw = 
  read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = 'Mr. Trash Wheel', range = 'A2:N408') %>% 
  janitor::clean_names() 

TW_df = 
  TW_df_raw %>% 
  drop_na('dumpster') %>% 
  mutate(sports_balls = round(sports_balls)) %>% 
  relocate(dumpster, year, month, date)

head(TW_df) %>% knitr::kable()
```

## Read and clean precipitation data for 2018 and 2019
```{r Read and clean precipitation data}
# read 2018 and 2019 data seperately
prec2018_df = 
  read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = '2018 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  drop_na()

prec2019_df = 
  read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = '2019 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  drop_na()

# combine precipitation datasets
prec_df = 
  bind_rows(prec2018_df, prec2019_df) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)

head(prec_df) %>% knitr::kable()
```

## Write a paragraph about these data
```{r, include = FALSE}
nrow(prec_df)
ncol(prec_df)
sum(pull(prec2018_df, total))
median(TW_df %>% 
      filter(year == 2017) %>% 
  pull(sports_balls))

median(TW_df_raw %>% 
      filter(year == 2017) %>% 
  pull(sports_balls))

```

In Mr.Trash Wheel dataset there are **`r nrow(TW_df)`** observations and **`r ncol(TW_df)`** features. In precipitation dataset there are **`r nrow(prec_df)`** observations and **`r ncol(prec_df)`** features, including **`r nrow(prec2018_df)`** months in 2018 and **`r nrow(prec2019_df)`** months in 2019.

In Mr.Trash Wheel dataset, *dumpster* is the count of dump time, ranging from `r min(pull(TW_df, dumpster))` to `r max(pull(TW_df, dumpster))`; *year*, *month* and *date* are time-related variables. Others are numeric variables including *weight_tons*, *volume_cubic_yards* and variables that are counts of objects. These objects include glass bottles, sport balls and grocery bags. In precipitation dataset, *year* and *month* are 
time-related variables. *Total* means total inches of precipitation that month, it ranges from `r min(pull(prec_df, total))` to `r max(pull(prec_df, total))` with mean precipitation equals to `r round(mean(pull(prec_df, total)),2)`.
 
For available data, the total precipitation in 2018 is **`r sum(pull(prec2018_df, total))`**. The median number of sports balls in a dumpster in 2017 is **`r median(TW_df %>% filter(year == 2017) %>% pull(sports_balls))`**. P.S. Also check that in raw data, the median is still **`r median(TW_df_raw %>% filter(year == 2017) %>% pull(sports_balls))`**.

























