---
title: "p8105_hw2_zl3119"
author: "Zheyan"
date: "2021/9/30"
output: github_document
---

```{r import packages, include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem1
Read and clean the Mr. Trash Wheel sheet:

* specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

## Read and clean Mr.Trash Wheel Sheet

```{r Read and clean Mr.Trash Wheel,warning=FALSE}
TW_df_raw = 
  read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = 'Mr. Trash Wheel', range = 'A2:N408') %>% 
  janitor::clean_names() 

TW_df = 
  TW_df_raw %>% 
  drop_na('dumpster') %>% 
  mutate(sports_balls = round(sports_balls)) %>% 
  relocate(dumpster, year, month, date)

head(TW_df) %>% knitr::kable()
```

## Read and clean precipitation data for 2018 and 2019
```{r Read and clean precipitation data}
# read 2018 and 2019 data seperately
prec2018_df = 
  read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = '2018 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  drop_na()

prec2019_df = 
  read_excel('data/Trash-Wheel-Collection-Totals-8-6-19.xlsx', sheet = '2019 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  drop_na()

# combine precipitation datasets
prec_df = 
  bind_rows(prec2018_df, prec2019_df) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)

head(prec_df) %>% knitr::kable()
```

## Briefly introduce these data
```{r, include = FALSE}
nrow(prec_df)
ncol(prec_df)
sum(pull(prec2018_df, total))
median(TW_df %>% 
      filter(year == 2017) %>% 
  pull(sports_balls))

median(TW_df_raw %>% 
      filter(year == 2017) %>% 
  pull(sports_balls))

```

In Mr.Trash Wheel dataset there are **`r nrow(TW_df)`** observations and **`r ncol(TW_df)`** features. In precipitation dataset there are **`r nrow(prec_df)`** observations and **`r ncol(prec_df)`** features, including **`r nrow(prec2018_df)`** months in 2018 and **`r nrow(prec2019_df)`** months in 2019.

In Mr.Trash Wheel dataset, *dumpster* is the count of dump time, ranging from `r min(pull(TW_df, dumpster))` to `r max(pull(TW_df, dumpster))`; *year*, *month* and *date* are time-related variables. Others are numeric variables including *weight_tons*, *volume_cubic_yards* and variables that are counts of objects. These objects include glass bottles, sport balls and grocery bags. In precipitation dataset, *year* and *month* are 
time-related variables. *Total* means total inches of precipitation that month, it ranges from `r min(pull(prec_df, total))` to `r max(pull(prec_df, total))` with mean precipitation equals to `r round(mean(pull(prec_df, total)),2)`.
 
For available data, the total precipitation in 2018 is **`r sum(pull(prec2018_df, total))`**. The median number of sports balls in a dumpster in 2017 is **`r median(TW_df %>% filter(year == 2017) %>% pull(sports_balls))`**. P.S. Also check that in raw data, the median is still **`r median(TW_df_raw %>% filter(year == 2017) %>% pull(sports_balls))`**.

# Problem 2

This problem uses the FiveThirtyEight data; these data were gathered to create the interactive graphic on this page. In particular, we’ll use the data in pols-month.csv, unemployment.csv, and snp.csv. Our goal is to merge these into a single data frame using year and month as keys across datasets.

## Step 1

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r clean the data in pols-month, warning=FALSE}
# make_int = function(x)(as.integer(x))
date_list = c('year','month','day')

pols_df_raw = read_csv('data/fivethirtyeight_datasets/pols-month.csv', show_col_types = FALSE) %>% janitor::clean_names()

pols_df = 
  pols_df_raw %>% 
  separate(mon, sep = '-', into = date_list) %>% 
  mutate_at(date_list, as.integer) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(prez = ifelse(prez_gop == 1, 'gop', ifelse(prez_dem == 1, 'dem', 'others'))) %>% 
  select(-prez_gop, -prez_dem, -day) %>% 
  relocate(year, month, prez)

head(pols_df) %>% knitr::kable()
```

## Step 2

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r clean the data in snp, warning=FALSE}
# try reading data in Excel format (cannot read csv file)
snp_df_raw = read_excel('data/fivethirtyeight_datasets/snp.xlsx') %>% janitor::clean_names()
date_list2 = c('month','day','year')

snp_df = 
  snp_df_raw %>% 
  separate(date, sep = '/', into = date_list2) %>% 
  mutate_at(date_list, as.integer) %>% 
  mutate(month = month.name[month]) %>%
  select(-day) %>% 
  relocate(year, month)

head(snp_df) %>% knitr::kable()
```

## Step 3

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r unemployment data}
# try reading data in Excel format (cannot read csv file)
unemploy_df_raw = read_excel('data/fivethirtyeight_datasets/unemployment.xlsx') %>% janitor::clean_names()

# temp month df to help convert month in unemployment data
temp_month_df = 
  tibble(month = substr(month.name, 1, 3),
         month_longer = month.name)

# tidy data
unemploy_df = 
  pivot_longer(
    unemploy_df_raw, 
    jan:dec,
    names_to = 'month',
    values_to = 'unemployment_rate') %>% 
  # Get Month variable uppercase for first letter to match
  mutate(month = str_to_title(month)) %>% 
  # Get the longer str of month
  left_join(temp_month_df, by = 'month') %>% 
  mutate(month = month_longer) %>% 
  select(-month_longer) %>% 
  # Also convert year to int to match the temp_month_df
  mutate(year = as.integer(year)) 

head(unemploy_df) %>% knitr::kable()
```

Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r Merge all the data}
# merging snp into pols
result_df = 
  left_join(pols_df, snp_df, by = c('year', 'month')) %>% 
  left_join(unemploy_df, by = c('year', 'month'))

# showing more rows this time
head(result_df, 15) %>% knitr::kable()
```

## Briefly introduce these data
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).


**pols_df** contains `r nrow(pols_df)` oberservations and `r ncol(pols_df)` variables related to the number of national politicians who are democratic or republican at each month. Years in the dataset ranges from `r min(pull(pols_df, year))` to `r max(pull(pols_df, year))`. The key president variable *prez* indicates the president is democratic or republican.

**snp_df** contains `r nrow(snp_df)` oberservations and `r ncol(snp_df)`  variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole.  Years in the dataset ranges from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. The key president variable *close* is the closing values of the S&P stock index on the associated date.

**unemploy_df** contains `r nrow(unemploy_df)` oberservations and `r ncol(unemploy_df)`  variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole.  Years in the dataset ranges from `r min(pull(unemploy_df, year))` to `r max(pull(unemploy_df, year))`. The key president variable *unemployment_rate* is percentage of unemployment in each month.

Since the range of years in the above 3 datasets vary, merge them togther into **result_df** with lefj-join can result in some variables with missing data.  In the final data, *close* contains `r sum(!complete.cases(pull(result_df, close)))` missing values. Also, *unemployment_rate*  contains `r sum(!complete.cases(pull(result_df, unemployment_rate)))` missing values

# Problem 3

This problem uses data from NYC Open data on the popularity of baby names.

## Load and tidy the data

Note that, although these data may seem fairly well formatted initially, the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will need to be removed

```{r tidy data of baby names}
babynames_df_raw = 
  read_csv('data/Popular_Baby_Names.csv') %>% janitor::clean_names()

babynames_df = 
  babynames_df_raw %>% 
  # drop duplicates
  distinct() %>% 
  # make first letter uppercase and the rest lower case
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  # remove rank (rank later)
  select(-rank) %>% 
  # group by each name and calcuate overall count
  group_by(year_of_birth, gender, ethnicity, childs_first_name) %>%
  summarize(overall_count = sum(count, na.rm = TRUE)) %>% 
  # give a new rank by year_of_birth, gender, ethnicity
  arrange(year_of_birth, gender, ethnicity, -overall_count) %>%
  group_by(year_of_birth, gender, ethnicity) %>% 
  # use rank() and min method to tie 
  mutate(rank = rank(-overall_count, ties.method = 'min'))

head(babynames_df, 10) %>% knitr::kable()
```


## Popularity of the name “Olivia”
Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.


## children with a name (y axis) against the rank in popularity of that nam
Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).














